---
title: "Harvest traits from EUCLID dump"
author: "AC"
date: "06/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

librarian::shelf(tidyverse, conflicted, DesiQuintans/desiderata)

conflict_prefer("pluck", "purrr")
conflict_prefer("filter", "dplyr")
```

# Aim

In `1_clean_EUCLID_dump.Rmd`, I processed the EUCLID Corymbia fact sheets that I downloaded so that each relevant section of the fact sheet was in a separate column, which makes it easier to target, for example, leaf traits without catching seed traits.

In this document, I will do the actual harvesting.


## Traits to harvest

- Classification
    - Subgenus
    - Section
    - Subsection
    - Series
    - Species
    - Subspecies
    - Victoria and Anne-Cecile: "For some species there is not all the 
      information and it does not indicate which information is what, for 
      instance in E. diversicolor appears as "Eucalyptus | Symphyomyrtus | 
      Inclusae" in which Inclusae is a subseries but in most appears as 
      "Eucalyptus | Symphyomyrtus | Maidenaria | Euryotae | Bridgesiana" being 
      second subgenus, third section, fourth subsection and final serie"

- General description
    - Habit
    - Maximum height
    - Regeneration strategy (comes after habit and maximum height in desc)
    
- Adult leaves
    - Blade form
    - Blade length
    - Blade width
    - Oil glands (intersectional, island, absent, or obscure, sometimes 
      'abundant' or 'crowded')
    - Reticulation of veins ("moderately to densely reticulate" or 
      "sparsely to moderately reticulate", for example)
    
- Fruit
    - Fruit length
    - Fruit width
    
- Seed
    - Seed length
    - Seed shape
    
- Inflorescence
    - Flower colour (useful for pilot analysis)



# 01. Import data

```{r}
raw <- 
    read_rds("../intermediate/factsheet_as_sections.rds") %>% 
    set_names(c("name", "classif", "form", "leaf", "fruit", "seed", 
                "flower", "auth", "desc")) %>% 
    glimpse()
```


# 02. Extract classification names

Genus, subgenus, section, subsection, series, and subseries can come from `classif`. Correctly sorting the classification names into columns will be challenging. May need to make a list of allowed names in each column?

Species and subspecies can come from `name`.

## List of allowed classification names

I retrieved this list using regular expressions from Dean Nicolle's *Classification of the Eucalypts* table (`raw_data/Classification-Of-The-Eucalypts.pdf`). I omitted these unpublished names:


```{r}
genus <- 
    c("Corymbia")

subgenus <- 
    c("Corymbia", "Blakella")

section <- 
    c("Calophyllae", "Corymbia", "Naviculares", "Torellianae", "Maculatae",
      "Abbreviatae")

series <- 
    c("Cymbiformes", "Jacobsianae", "Trachyphloiae", "Dorsiventrales", "Isobilaterales",
      "Terminalipterae", "Tessellatae", "Scutiformes")

subseries <- 
    c("Pletiformes", "Limitaneae")
```

## Annotate names with their classification level

There's not that many unique classification paths (`r howmany(raw$classif)`). It might be best to just manually work my way through the paths using the above lists.

```{r}
class_paths <- 
    unique(raw$classif) %>% 
    sort() %>% 
    # Prefix each name with a hyphen. Names starting with a hyphen haven't been
    # matched yet. Later, names that are matched will have the hyphen removed.
    # str_replace_all("([A-Z])", "-\\1") %>% 
    # Split the paths using the bar. It's now a nested list.
    str_split(fixed(" | "), n = Inf, simplify = FALSE)

head(class_paths)


annotate_path <- function(path, ...) {
    # `path` is a vector of the plant's classification names, starting from 
    # genus and going down to subseries. All 6 names are rarely present.
    
    # `namelist` is a list of namelists. [[1]] is genus, [[2]] is subgenus, all 
    # the way down to [[6]] which is subseries.
    namelist <- list(...)
    
    # `level` is the name of each classification level, derived from the 
    # variable names. This will be used to annotate the path.
    levels <- substitute(list(...)) %>% as.character() %>% {.[-1]}
    
    
    
    # The idea is to iterate through each element in `path` by stepping through
    # the lists in `names`. If a match is found, annotate the path with 
    # 1. For path[1], iterate through names[1:6].
    # 2. When a match is found, add level[n] to the front of path[1].
    # 3. If nothing is found, add "unknown".
    # 4. For path[2], iterate through names[2:6].
    # 5. Repeat 2:3.
    output <- character(0)
    
    for(i in seq_along(path)) {
        # For each item in the plant's classification path...
        item <- path[[i]]
        
        # Choose the lists to compare it the item to. When the second path item is
        # being compared (should correspond to subgenus), the first namelist (genus)
        # is not needed.
        for(j in i:length(namelist)) {
            if (item %in% namelist[[j]]) {
                output <- append(output, paste(levels[[j]], item))
                break()
            }
        }
    }
    
    output <- str_flatten(output, collapse = " | ")
    
    # If output is still empty, then no matches were found.
    return(ifelse(length(output) == 0, NA, output))
}

named_class_paths <- 
    map_chr(class_paths, ~ annotate_path(., genus, subgenus, section, 
                                        series, subseries))
    

head(named_class_paths)
```

These annotated paths can then be joined back into the original table.

```{r}
classif_table <- 
    tibble(classif = unique(raw$classif) %>% sort(),
           named_classif = named_class_paths) %>% 
    glimpse()

raw_classification <- 
    raw %>% 
    select(name, classif) %>% 
    left_join(classif_table, by = "classif")
```


## Split names into columns

```{r}
get_level <- function(path, level) {
    # The x ? part is to catch hybrids, e.g. Eucalyptus x phylacis, where 
    # "x phylacis" is an accepted species name.
    out <- str_match(path, paste(level, "((x )?\\w+)"))[2]
}

classification_df <- 
    raw_classification %>% 
    group_by(name) %>% 
    mutate(genus = word(name, 1) %>% str_to_title(),
            subgenus = get_level(named_classif, "subgenus"),
            section = get_level(named_classif, "section"),
            series = get_level(named_classif, "series"),
            subseries = get_level(named_classif, "subseries"),
            species = get_level(name, "^eucalyptus"),
            subspecies = get_level(name, "subsp\\."),
            variety = get_level(name, "var\\.")) %>% 
    ungroup() %>% 
    select(-classif, -named_classif) %>% 
    glimpse()
```



# 03. Extract general growth details

Habit is sometimes the first word, but not always. It's usually tree, mallee (sometimes "Mallet"). Sometimes it is both. TRUE/FALSE column?


## Retrieve habit

```{r}
word(raw$form, 1, 2) %>% 
    str_split(" ", simplify = TRUE) %>% 
    count_unique(sort = TRUE)
```

Given this, there are only a few growth types in the descriptions.

```{r}
tree_regex <- regex("tree", ignore_case = TRUE)
mallee_regex <- regex("malle", ignore_case = TRUE)
# "Unlike the mallee, [the marlock] is single-stemmed and lacks a lignotuber."
marlock_regex <- regex("marlock", ignore_case = TRUE)
shrub_regex <- regex("shrub", ignore_case = TRUE)

habit_df <- 
    raw %>% 
    select(name, form) %>% 
    group_by(name) %>% 
    mutate(habit_tree    = str_detect(form, tree_regex),
           habit_mallee  = str_detect(form, mallee_regex),
           habit_marlock = str_detect(form, marlock_regex),
           habit_shrub   = str_detect(form, shrub_regex)) %>% 
    # If all are false, then replace with NA.
    mutate(logical_sum = sum(habit_tree, habit_mallee, habit_marlock, 
                             habit_shrub, na.rm = TRUE)) %>% 
    mutate_at(vars(habit_tree:habit_shrub), 
              ~ ifelse(logical_sum == 0, NA, .)) %>% 
    ungroup() %>% 
    select(-form, -logical_sum) %>% 
    glimpse()

# Habit was retrieved for everything!
rows_with_na(habit_df)
```


## Retrieve maximum height

Maximum height will be the largest number in this column.

```{r}
# Retrieve numbers from a string. Use `fun` to choose which number to 
# return, e.g. max or min or median.
fetch_number <- function(str, fun) {
    nums <- 
        str %>% 
        str_split(" |-", simplify = TRUE) %>% 
        {shush(parse_number(.))} %>% 
        na_rm()
    
    if (length(nums) == 0) {
        return(NA)
    } else {
        do.call(fun, list(nums))
    }
}

height_df <- 
    raw %>% 
    select(name, form) %>% 
    group_by(name) %>% 
    mutate(height_max_m = fetch_number(form, max)) %>% 
    ungroup() %>% 
    select(-form) %>%
    glimpse()


# These NA rows were checked and confirmed to have no described height.
rows_with_na(height_df)
```


## Retrieve regeneration strategy

Regeneration strategy is at the end of this column's sentence. Nearly all mention a lignotuber, but sometimes it is to say that one is present or absent. 9 species don't mention one at all.

```{r}
raw$form %>% 
    str_split("\\. ") %>% 
    map(., ~ .[length(.)]) %>% 
    unlist() %>% 
    count_unique(sort = TRUE)
```

```{r}
# This list should be searched first because of the annoying phrasing of 
# "Not forming a lignotuber", which collides with the has_lignotuber list.
no_lignotuber <- c("Lignotuber absent",
                   "Non-lignotuberous",
                   "Lacking a lignotuber",
                   "lacking a lignotuber",
                   "Lignotuber apparently not formed",
                   "Lignotuber not recorded but expected to be absent",
                   "Lignotuber usually absent",
                   "Not forming a lignotuber")

has_lignotuber <- c("Forming a lignotuber",
                    "forming a lignotuber",
                    "Lignotuber present",
                    "Forms a lignotuber",
                    "Lignotuberous",
                    "seem to form a lignotuber",
                    "forming a \\?lignotuber")

lignotuber_varies <- c("Lignotuber present or absent",
                       "Trees with lignotubers and others lacking")

lignotuber_unknown <- c("Lignotuber character unknown",
                        "Lignotuber presence/absence unknown",
                        "Lignotuber status unknown",
                        "Presence or absence of lignotuber uncertain")

resprout_epicormic <- c("epicormic buds")

resprout_latroot <- c("lateral roots")


regeneration_df <- 
    raw %>% 
    select(name, form) %>% 
    group_by(name) %>% 
    mutate(regen_lignotuber = 
               case_when(any(str_detect(form, no_lignotuber))      ~ FALSE,
                         any(str_detect(form, lignotuber_varies))  ~ TRUE,
                         any(str_detect(form, has_lignotuber))     ~ TRUE,
                         any(str_detect(form, lignotuber_unknown)) ~ NA,
                                                              TRUE ~ NA),
           regen_epi_buds  = any(str_detect(form, resprout_epicormic)),
           regen_lat_roots = any(str_detect(form, resprout_latroot)),
           ) %>% 
    ungroup() %>% 
    select(-form) %>%
    glimpse()

# These NA rows were confirmed to have no description re. lignotubers.
rows_with_na(regeneration_df)
```



# 04. Extract adult leaf information


## Retrieve leaf blade shape

Blade form usually comes after the word "blade", but of course not always lol.

```{r}
# From https://en.wikipedia.org/wiki/Glossary_of_leaf_morphology
# Sorted from longest word to shortest, because short words are more likely to 
# be nested inside other names and should be tested last, e.g. 'ovate' 
# inside 'obovate'.
blade_shapes <- 
    c("palmatipartite", "pinnatipartite", "asymmetrical", "oblanceolate", 
      "pinnatisect", "spear-shaped", "palmatifid", "palmatisect", "pinnatifid", 
      "auriculate", "emarginate", "fenestrate", "flabellate", "lanceolate", 
      "mucronate", "obtrullate", "pandurate", "perfoliate", "semiterete", 
      "unifoliate", "acuminate", "apiculate", "attenuate", "cordiform", 
      "cuspidate", "laciniate", "obcordate", "orbicular", "perforate", 
      "sagittate", "spatulate", "subobtuse", "trullate", "acicular", 
      "aristate", "caudate", "digitate", "elliptic", "ensiform", "filiform", 
      "multifid", "reniform", "rhomboid", "subulate", "truncate", "undulate", 
      "cordate", "cuneate", "deltate", "deltoid", "falcate", "hastate", 
      "laminar", "oblique", "obovate", "palmate", "pedate", "peltate", 
      "plicate", "retuse", "rhombic", "rounded", "sinuate", "linear", "lorate", 
      "lyrate", "oblong", "obtuse", "terete", "acute", "lobed", "ovate") %>% 
    paste(collapse = "|") %>% 
    paste0("(", ., ")") %>% 
    regex(ignore_case = TRUE)

get_blade_shape <- function(str) {
    out <- str_extract(str, "blade .*?,")

    if(is.na(out)) {
        out <- str_extract(str, "^.*?(wide|base)")
    }
    
    return(out)
}

raw$leaf %>% 
    map_chr(., ~ get_blade_shape(.)) %>% 
    str_extract_all(blade_shapes) %>%
    unlist() %>% 
    count_unique(sort = TRUE)
```

```{r}
blade_df <- 
    raw %>% 
    select(name, leaf) %>% 
    group_by(name) %>% 
    # mutate(blade = str_extract(leaf, "blade .*?,")) %>% 
    mutate(blade = get_blade_shape(leaf)) %>% 
    mutate(blade_lanceolate = str_detect(leaf, "lanceolate"),
           blade_falcate = str_detect(leaf, "falcate"),
           blade_elliptic = str_detect(leaf, "elliptic"),
           blade_linear = str_detect(leaf, "linear"),
           blade_ovate = str_detect(leaf, "ovate"),
           blade_oblong = str_detect(leaf, "oblong"),
           blade_orbicular = str_detect(leaf, "orbicular"),
           blade_cordate = str_detect(leaf, "cordate"),
           blade_deltoid = str_detect(leaf, "deltoid"),
           blade_oblanceolate = str_detect(leaf, "oblanceolate"),
           blade_obovate = str_detect(leaf, "obovate"),
           blade_obcordate = str_detect(leaf, "obcordate")) %>% 
    ungroup() %>% 
    select(-leaf, -blade) %>%
    glimpse()

# Blade shapes were retrieved for all species!
rows_with_na(blade_df)
```


## Retrieve blade dimensions

Length and width comes after the blade form, and before the word "base".

There are actually only 2 sets of measurements in the leaf section, petiole length and blade length/width. If I remove petiole length (which I was not asked to supply), then I can just search the blade width/length manually. There are also numbers in brackets (e.g. "5.5--8(10) cm long", maybe rare measurements?) that I'd like to ignore.

```{r}
raw$leaf %>% 
    # Petiole seems to always come before 'blade', but I will remove it two 
    # different ways to be sure.
    str_extract("blade.*?$") %>% 
    str_remove("petiole.*?long") %>%
    str_extract_all("\\d.*?(long|wide)") %>% 
    unlist() %>% 
    str_remove_all("(^|\\()[0-9.]+\\)") %>% 
    head()
```

```{r}
as_mm <- function(val, unit) {
    if(is.na(val) | is.na(unit)) {
        return(NA_real_)
    }
    
    num <- parse_number(as.character(val))
    
    out <- case_when(unit == "mm" ~ num * 1,
                     unit == "cm" ~ num * 10,
                     unit == "m"  ~ num * 1000,
                             TRUE ~ NA_real_)
    return(out)
}

blade_dims_df <- 
    raw %>% 
    select(name, leaf) %>% 
    group_by(name) %>% 
    mutate(blade = 
               str_remove(leaf, "petioles? (to )?[0-9\\-\\.]{1,20} cm long") %>%
               str_remove_all("(^|\\()[0-9.]+\\)") %>% 
               str_replace_all("cm", " cm") %>% 
               str_replace_all("(\\d) ?--? ?(\\d)", "\\1--\\2") %>% 
               str_squish()) %>% 
    mutate(leaf_length = str_match(blade, "([0-9\\-\\.]+ \\w+) long")[2],
           leaf_width  = str_match(blade, "([0-9\\-\\.]+ \\w+) wide")[2]) %>%
    separate(col  = leaf_length, 
             into = c("leaf_length_min", "leaf_length_max", "leaf_length_unit"),
             sep  = "(--| )", fill = "left", remove = FALSE) %>% 
    separate(col  = leaf_width, 
             into = c("leaf_width_min", "leaf_width_max", "leaf_width_unit"),
             sep  = "(--| )", fill = "left", remove = FALSE) %>% 
    mutate(leaf_length_max = as_mm(leaf_length_max, leaf_length_unit),
           leaf_length_min = as_mm(leaf_length_min, leaf_length_unit),
           leaf_width_max  = as_mm(leaf_width_max,  leaf_width_unit),
           leaf_width_min  = as_mm(leaf_width_min,  leaf_width_unit)) %>% 
    select(-ends_with("_unit")) %>% 
    rename_all(~ str_replace(., "_(min|max)", "_mm_\\1")) %>% 
    ungroup() %>% 
    select(-leaf, -blade, -leaf_length, -leaf_width) %>%
    glimpse()

# These rows were checked and confirmed to be missing descriptions of one or 
# both leaf dimensions 
rows_with_na(blade_dims_df)
```


## Retrieve vein reticulation

Reticulation is near the words "reticulation" or "reticulate". Unfortunately, those words aren't always used.

```{r}
fetch_reticulation <- function(str) {
    # Fetch the fragment of the sentence that talks about reticulation
    out <- str_extract(str, ",[^,]*?(reticulat)[[:print:]]*?(,|\\.|$)")
    
    # Sometimes the description talks about venation instead, but reticulation
    # should be preferred.
    if(is.na(out)) {
        out <- str_extract(str, ",[^,]*?(venation)[[:print:]]*?(,|\\.|$)")
    }
    
    out %>% 
        str_remove_all(",") %>% 
        str_squish()
}

raw$leaf %>% 
    fetch_reticulation() %>% 
    count_unique(sort = TRUE) %>% 
    pull("unique")

# The words 'broken', 'clear', and 'finely' are used as descriptors sometimes 
# ("very dense and broken" or "reticulate and fine"). I omit these words.
sparse_retic     <- c("sparsely", "sparse")
moderate_retic   <- c("moderately", "moderate")
dense_retic      <- c("densely", "dense")
very_dense_retic <- c("very densely", "very dense")
obscure_retic    <- c("obscure")
unknown_retic    <- c("uncertain", "unknown")
absent_retic     <- c("absent", "invisible", "not visible", "no reticulation")
```

```{r}
venation_df <- 
    raw %>% 
    select(name, leaf) %>% 
    group_by(name) %>% 
    mutate(veins = fetch_reticulation(leaf)) %>% 
    mutate(reticulation_sparse = any(str_detect(veins, sparse_retic)),
           reticulation_moderate = any(str_detect(veins, moderate_retic)),
           reticulation_dense = any(str_detect(veins, dense_retic)),
           reticulation_very_dense = any(str_detect(veins, very_dense_retic)),
           reticulation_obscure = any(str_detect(veins, obscure_retic)),
           reticulation_unknown = any(str_detect(veins, unknown_retic)),
           reticulation_absent = any(str_detect(veins, absent_retic))
           ) %>% 
    # I manually checked the rows with NA venation, and only E. kybeanensis
    # needs to be fixed.
    mutate_at(vars(starts_with("reticulation")),
              ~ ifelse(name == "eucalyptus kybeanensis", FALSE, .)) %>%
    mutate(reticulation_absent = ifelse(name == "eucalyptus kybeanensis", 
                                        TRUE, reticulation_absent)) %>% 
    ungroup() %>% 
    select(-leaf, -veins) %>%
    glimpse()


# These rows were checked and confirmed to be missing venation descriptions.
rows_with_na(venation_df)
```


## Retrieve oil glands

Oil glands are near the end of the passage and can fall into several categories, so it might be another TRUE/FALSE set of columns.

```{r}
raw$leaf %>% 
    str_extract(",[^,]*?oil glands.*?$") %>%
    str_remove_all(",") %>% 
    str_squish() %>% 
    str_split(" ") %>% 
    unlist() %>% 
    str_remove_all("[[:punct:]]") %>% 
    count_unique(sort = TRUE) %>% 
    filter(count >= 10)

gland_words <- 
    c("island", "intersectional", "obscure", "absent")
```

```{r}
gland_df <- 
    raw %>% 
    select(name, leaf) %>% 
    group_by(name) %>% 
    mutate(gland = 
               str_extract(leaf, ",[^,]*?oil glands.*?$") %>%
               str_remove_all(",") %>% 
               str_squish()) %>% 
    mutate(oil_glands_island = str_detect(gland, "island"),
           oil_glands_intersectional = str_detect(gland, "intersectional"),
           oil_glands_obscure = str_detect(gland, "obscure"),
           oil_glands_absent = str_detect(gland, "absent"),
           ) %>% 
    ungroup() %>% 
    select(-leaf, -gland) %>%
    glimpse()

# These rows were checked and confirmed to be missing gland descriptions.
rows_with_na(gland_df)
```



# 05. Extract fruit dimensions

Fruit length and width are in the middle of the passage, but care must be taken to avoid the pedicel length. Fruit length and width are always mentioned in that order, but length is also not always present heh.

```{r}
fruit_dims_df <- 
    raw %>% 
    select(name, fruit) %>% 
    group_by(name) %>% 
    mutate(dims = 
               str_remove(fruit, "^.*?(hemispher|cup|cylind|obconic)") %>% 
               str_remove("pedicel.{0,25}long.*?(,|;)") %>%
               str_remove_all("(^|\\()[0-9.]+\\)") %>% 
               str_remove_all("--[0-9.]+--") %>% 
               str_replace_all("cm (long|wide)", " cm \\1") %>% 
               str_squish()) %>% 
    mutate(fruit_length = str_match(dims, "([0-9\\-\\.]+ \\w+) long")[2],
           fruit_width  = str_match(dims, "([0-9\\-\\.]+ \\w+) wide")[2]) %>%
    mutate(fruit_length = 
               ifelse(str_detect(dims, "and long"), fruit_width, fruit_length),
           fruit_width = 
               ifelse(str_detect(dims, "and wide"), fruit_length, fruit_width)) %>% 
    # Two species were checked and need to be manually fixed to deal with typos.
    mutate(fruit_length = ifelse(name == "eucalyptus misella", 
                                 "0.4--0.5 cm", fruit_length),
           fruit_width = case_when(name == "eucalyptus misella" ~ "0.6--0.8 cm",
                                   name == "eucalyptus nigrifunda" ~ "0.1--0.5 cm",
                                   TRUE ~ fruit_length)
          ) %>% 
    separate(col  = fruit_length, 
             into = c("fruit_length_min", "fruit_length_max", "fruit_length_unit"),
             sep  = "(--| )", fill = "left", remove = FALSE) %>% 
    separate(col  = fruit_width, 
             into = c("fruit_width_min", "fruit_width_max", "fruit_width_unit"),
             sep  = "(--| )", fill = "left", remove = FALSE) %>% 
    mutate(fruit_length_max = as_mm(fruit_length_max, fruit_length_unit),
           fruit_length_min = as_mm(fruit_length_min, fruit_length_unit),
           fruit_width_max  = as_mm(fruit_width_max,  fruit_width_unit),
           fruit_width_min  = as_mm(fruit_width_min,  fruit_width_unit)) %>% 
    select(-ends_with("_unit")) %>% 
    rename_all(~ str_replace(., "_(min|max)", "_mm_\\1")) %>% 
    ungroup() %>% 
    select(-fruit_length, -fruit_width) %>% 
    glimpse()


# These rows that contain NAs have been checked to confirm that they are indeed
# missing some min/max records.
rows_with_na(fruit_dims_df)
```



# 06. Extract seed information

## Retrieve seed length

Seed length is the only numeric value in this field.

```{r}
seed_dims_df <- 
    raw %>% 
    select(name, seed) %>% 
    group_by(name) %>% 
    mutate(dims = 
               str_replace_all(seed, "mm,", "mm long") %>% 
               str_replace_all("diameter", "long") %>% 
               str_remove_all("(^|\\()[0-9.]+\\)") %>% 
               str_replace_all("(\\d)-(\\d)", "\\1--\\2") %>% 
               str_remove_all("--[0-9.]+--") %>% 
               str_replace_all("mm long", " mm long") %>% 
               str_squish()) %>% 
    mutate(seed_length = str_match(dims, "([0-9\\-\\.]+ \\w+) long")[2]) %>%
    separate(col  = seed_length, 
             into = c("seed_length_min", "seed_length_max", "seed_length_unit"),
             sep  = "(--| )", fill = "left", remove = FALSE) %>% 
    mutate(seed_length_max = as_mm(seed_length_max, seed_length_unit),
           seed_length_min = as_mm(seed_length_min, seed_length_unit)) %>% 
    select(-ends_with("_unit")) %>% 
    rename_all(~ str_replace(., "_(min|max)", "_mm_\\1")) %>% 
    ungroup() %>% 
    select(-seed, -dims, -seed_length) %>%
    glimpse()


# These rows with NAs were checked and confirmed to be missing seed length data.
rows_with_na(seed_dims_df)
```


## Retrieve seed shape

Seed shape is the sentence fragment immediately after the seed length.

```{r}
raw$seed %>% 
    str_remove("^.*?(long|diameter|mm)(,|;|\\.) ?") %>% 
    str_remove("dorsal.*?$") %>% 
    str_extract("^.*?,") %>%
    str_remove_all(",") %>% 
    unlist() %>% 
    count_unique(sort = TRUE) %>% 
    filter(count >= 6)

seed_shape_regex <- 
    c("obliquely-pyramidal", "obliquely pyramidal", "flattened-ovoid", 
      "flattened ovoid", "sub-spherical", "pyramidal", "cuboid", "ovoid") %>% 
    paste(collapse = "|") %>% 
    paste0("(", ., ")") %>% 
    regex(ignore_case = TRUE)

raw$seed %>% 
    str_extract_all(seed_shape_regex) %>% 
    unlist() %>% 
    count_unique(sort = TRUE)
```

```{r}
seed_shape_df <- 
    raw %>% 
    select(name, seed) %>% 
    group_by(name) %>% 
    mutate(shape = 
               seed %>% 
               str_extract_all(seed_shape_regex) %>% 
               unlist() %>% 
               paste0(collapse = ";") %>% 
               paste0(";", ., ";")) %>% 
    mutate(seed_shape_flat_ovoid        = str_detect(shape, ";flattened.ovoid;"),
           seed_shape_ovoid             = str_detect(shape, ";ovoid;"),
           seed_shape_oblique_pyramidal = str_detect(shape, ";obliquely.pyramidal;"),
           seed_shape_pyramidal         = str_detect(shape, ";pyramidal;"),
           seed_shape_cuboid            = str_detect(shape, ";cuboid;"),
           seed_shape_subspherical      = str_detect(shape, ";sub-spherical;")
           ) %>% 
    # If all are false, then replace with NA.
    mutate(logical_sum = sum(seed_shape_flat_ovoid, seed_shape_ovoid,
                             seed_shape_oblique_pyramidal, seed_shape_pyramidal,
                             seed_shape_cuboid, seed_shape_subspherical,
                             na.rm = TRUE)) %>% 
    mutate_at(vars(seed_shape_flat_ovoid:seed_shape_subspherical), 
              ~ ifelse(logical_sum == 0, NA, .)) %>% 
    ungroup() %>% 
    select(-seed, -shape, -logical_sum) %>% 
    glimpse()

# The species here were checked and do not use standard terms for seed shape, or 
# they use rare words. These include "linear", "boat-shaped", "saucer-shaped",
# "flattened", and "elongated".
rows_with_na(seed_shape_df)
```



# 07. Extract flower colour

Flower colour is the very last sentence, usually the very last word. Species often have variable flower colours, so it is best to record each colour in a separate column. The users can simplify this as desired.


```{r}
raw$flower %>% 
    str_match(" Flowers (.*?)\\.$") %>% 
    {.[,2]} %>% 
    str_split(boundary("word")) %>% 
    unlist() %>% 
    count_unique(sort = TRUE)

col_white        <- c("white", "cream", "whitish")
col_yellow       <- c("yellow(ish| |\\.|$)", "lemon")
col_red          <- c("red", "pink", "pink-red", "pinkish", "reddish")
col_green_yellow <- c("green", "yellow-green", "green-yellow")
col_orange       <- c("orange")
col_purple       <- c("purple", "mauve")
col_missing      <- c("Flowers not seen.", "unknown", "not known")
```

The colours need to be drawn specifically from the last sentence, or else there will be unwanted matches.

```{r}
first_n_words <- function(str, start, end) {
    words <- str_split(str, " ")
    
    map_chr(words, 
        function(x) {
            limit <- ifelse(end > length(x), length(x), end)
            
            return(paste(x[start:limit], collapse = " "))
        })
}

inflorescence_df <- 
    raw %>% 
    select(name, flower) %>% 
    group_by(name) %>% 
    mutate(colour = 
               str_extract(flower, "Flowers? .*?\\.?$") %>% 
               str_remove_all("rarely.*?$") %>% 
               str_remove_all("pale|usually|or |to") %>% 
               str_squish() %>% 
               first_n_words(1, 5),
           flower_white = any(str_detect(colour, col_white)),
           flower_yellow = any(str_detect(colour, col_yellow)),
           flower_green_yellow = any(str_detect(colour, col_green_yellow)),
           flower_red = any(str_detect(colour, col_red)),
           flower_orange = any(str_detect(colour, col_orange)),
           flower_purple = any(str_detect(colour, col_purple))) %>% 
    # If no columns were filled in, replace them with NAs.
    mutate(missing = sum(flower_white, flower_yellow, flower_green_yellow, 
               flower_red, flower_orange, flower_purple, na.rm = TRUE)) %>% 
    mutate_at(vars(flower_white:flower_purple),
              ~ ifelse(missing == 0, NA, .)) %>%
    ungroup() %>% 
    select(-flower, -colour, -missing) %>%
    glimpse()


# These NA rows were confirmed to have no colour description.
rows_with_na(inflorescence_df)
```



# 08. Combine result dataframes

```{r}
glimpse(classification_df)
glimpse(habit_df)
glimpse(height_df)
glimpse(regeneration_df)
glimpse(blade_df)
glimpse(blade_dims_df)
glimpse(venation_df)
glimpse(gland_df)
glimpse(seed_dims_df)
glimpse(seed_shape_df)
glimpse(inflorescence_df)

corymbia_traits <-  
    classification_df %>% 
    left_join(habit_df, by = "name") %>% 
    left_join(height_df, by = "name") %>% 
    left_join(regeneration_df, by = "name") %>% 
    left_join(blade_df, by = "name") %>% 
    left_join(blade_dims_df, by = "name") %>% 
    left_join(venation_df, by = "name") %>% 
    left_join(gland_df, by = "name") %>% 
    left_join(seed_dims_df, by = "name") %>% 
    left_join(seed_shape_df, by = "name") %>% 
    left_join(inflorescence_df, by = "name") %>% 
    left_join(select(raw, name, botanical_desc = desc), by = "name")
```


## Save output

```{r}
write_csv(corymbia_traits, "../outputs/corymbia_traits_from_EUCLID.csv")
write_rds(corymbia_traits, "../outputs/corymbia_traits_from_EUCLID.rds")
```


## Data dictionary

```{r}
glimpse(corymbia_traits)
```

- **name** --- Unique ID column.
- **genus** to **variety** --- The species' taxonomic names.
- **habit_tree** to **habit_shrub** --- The species' growth habit.
- **height_max_m** --- The species' maximum height in metres.
- **regen_lignotuber** to **regen_lat_roots** --- The species' regeneration strategy, whether lingotubers, reshooting from epicormic shoots, or reshooting from lateral roots.
- **blade_lanceolate** to **blade_obcordate** --- The shape of adult leaves.
- **leaf_length_mm_min** to **leaf_width_mm_max** --- The minimum and maximum dimensions of adult leaves in millimetres.
- **reticulation_sparse** to **reticulation_absent** --- The reticulate venation of adult leaves. 
- **oil_glands_island** to **oil_glands_absent** --- The arrangement of oil glands on adult leaves.
- **seed_length_mm_min** to **seed_length_mm_max** --- The dimensions of seeds.
- **seed_shape_flat_ovoid** to **seed_shape_subspherical** --- The shape of seeds.
- **flower_white** to **flower_purple** --- The recorded colours of the filaments.
- **botanical_desc** --- The complete botanical description. This can make the dataframe lag badly, so remove it if not needed.

